### 2023-06-16 appendix.tex
up

Reset VASS extend the basic VASS model with special “reset
transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
\begin{definition}
A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
% $\mathds{Z}^d$
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
%$T \subseteq Q \times \mathds{Z}^d \times Q$
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

@@

\section{Reset VASS and state-resilience}\label{appendix}

@@

The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS%, whose proof can be seen in Appendix~\ref{appendix}%
.
\begin{proposition}\label{liveness reset}
The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
\end{proposition}

@@

%\begin{proposition}\label{liveness reset}
% The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
%\end{proposition}

>>>>

Reset-VASS extend the basic VASS model with special “reset
transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
\begin{definition}
A {\em reset-VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
% $\mathds{Z}^d$
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
%$T \subseteq Q \times \mathds{Z}^d \times Q$
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

@@

\section{Reset-VASS and state-resilience}\label{appendix}

@@

The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS%, whose proof can be seen in Appendix~\ref{appendix}%
.
\begin{proposition}\label{liveness reset}
The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
\end{proposition}

@@

%\begin{proposition}\label{liveness reset}
% The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
%\end{proposition}

### 2023-06-16 state-resilience.tex
up

We start with the case $\Safe = \uparrow \Safe$.
%
Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
This stems from the fact that it is undecidable in the particular case of Reset VASS,
where  $t$-liveness is both undecidable and 
 reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

@@

\begin{proof}
{\sc State-resilience} itself is undecidable % from Proposition~\ref{reductions} 
since 
{\sc $t$-liveness} is undecidable in Reset VASS
and reducible to {\sc State-resilience}.
Additionally, in WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
reducible to {\sc $k$-state-resilience}:
since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
and there exists $n_0 \in \N$ such that 
$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
We compute 
$n_0$, then iteratively check whether $k$-state-resilience 
hold for $k$ from $0$ to $n_0$.  
Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
since 
$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
Hence the undecidability of {\sc Bounded-state-resilience}
and 
 {\sc $k$-state-resilience}. 
\end{proof}

@@

More concretely,
VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
It is well-known that 
VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
Hence {\sc State-resilience} is decidable for %both Petri nets and 
VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:

@@

\begin{proof}
Let us show that Reset VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. 
Then, {\sc State-resilience} would be decidable for Reset VASS, which leads to a contradiction.
\end{proof}

Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience }. Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889}:

@@

Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$. As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by Reset-VASS, as seen in Appendix~\ref{HS-Minsk}.

>>>>

We start with the case $\Safe = \uparrow \Safe$.
%
Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
This stems from the fact that it is undecidable in the particular case of reset-VASS,
where  $t$-liveness is both undecidable and 
 reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

@@

\begin{proof}
{\sc State-resilience} itself is undecidable % from Proposition~\ref{reductions} 
since 
{\sc $t$-liveness} is undecidable in reset-VASS
and reducible to {\sc State-resilience}.
Additionally, in WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
reducible to {\sc $k$-state-resilience}:
since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
and there exists $n_0 \in \N$ such that 
$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
We compute 
$n_0$, then iteratively check whether $k$-state-resilience 
hold for $k$ from $0$ to $n_0$.  
Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
since 
$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
Hence the undecidability of {\sc Bounded-state-resilience}
and 
 {\sc $k$-state-resilience}. 
\end{proof}

@@

More concretely,
VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
It is well-known that 
VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
Hence {\sc State-resilience} is decidable for %both Petri nets and 
VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  \iffalse Moreover, we may show:

@@

\begin{proof}
Let us show that reset-VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for reset-VASS. 
Then, {\sc State-resilience} would be decidable for reset-VASS, which leads to a contradiction.
\end{proof} \fi
Moreover there exist classes of WSTSs with strong 
 compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$, such as reset-VASS.

Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience}. Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:

@@

We now consider the case $\Safe = \downarrow \Safe$.
%
Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$. As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
